

This PhD journey has been long, challenging, and at times exhausting, but it has also been filled with rewarding moments and incredible people who have made it all worthwhile.

First and foremost, I want to express my deepest gratitude to my supervisor, Laurent Barillé. Thank you for your guidance, patience, and unwavering support throughout this process. Thank you for all the freedom you gave me during those years. Your expertise and encouragement pushed me to improve at every step, even when I doubted myself. I have learned so much under your mentorship, and I am truly grateful for your trust in me. Also, a special thanks for your help with all the administrative stuff—I know I’ve been lazy about it, and I truly appreciate your patience!

A huge thank you as well to my co-supervisor, Pierre Gernez, for your insightful advice, constructive feedback, and constant support. Your perspective has always helped me take a step back and see things more clearly, and I have deeply appreciated our discussions. 

I have been incredibly lucky to work alongside Bede Davies, whose generosity with his time and knowledge has been beyond what I could have hoped for in a colleague. Bede, collaborating with you has been one of the best parts of this PhD—your patience and dedication made even the toughest research challenges more manageable.

I also want to sincerely thank Philippe Rosa, who has been an absolute lifesaver during fieldwork. Philippe, you were always there, ready to help, problem-solve, or just share a laugh when things inevitably went wrong in the field.

On a more personal note, I owe an immense amount of gratitude to Laura Zoffoli. Thank you for putting up with the late nights, the stress, and the endless conversations about seagrass and remote sensing. Your support and patience have meant the world to me, and I couldn’t have done this without you.

Finally, to my family—thank you for always believing in me, even when I didn’t believe in myself. Your encouragement, love, and reminders that there is life outside of research have been a constant source of strength. I am so lucky to have you in my corner.

To everyone who has been part of this journey in one way or another—thank you.

